---
title: "Explore and Mine Data"
author: "Sahil Sah, Veniamin Knyazev"
output: html_notebook
---


# Introduction

This report presents the results of the following analytical queries against our MySQL data warehouse, aiming to provide insights into the publishing patterns of various journals:

1. Top five journals with the most articles published in them for the time period.
2. Number of articles per journal per year broken down by quarter.


# Setup for MySQL Instance

```{r setup, include=FALSE}
library(RMySQL)
library(DBI)
library(ggplot2)
# Establish a connection to the MySQL database
db_user <- 'admin'
db_password <- 'tik12tik'
db_name <- 'pubmedDB'
db_host <- 'practicum2.cxdn8klm55ui.us-east-1.rds.amazonaws.com'
db_port <- 3306

mysqlCon <-  dbConnect(MySQL(), user = db_user, password = db_password,
                   dbname = db_name, host = db_host, port = db_port)
```

## Analytical Query I: Top five journals with the most articles published in them for the time period.

```{r, top_5_journals, eval = T, warning = F}
# Store the result to a dataframe
top_5_journals <- dbGetQuery(mysqlCon, 
   "SELECT jd.title AS journal_title, SUM(jf.articles_count) AS sumOfArticles
    FROM Journal_Facts jf JOIN Journal_Dim jd ON jf.journal_id = jd.journal_id
    GROUP BY jd.title
    ORDER BY sumOfArticles DESC
    LIMIT 5")
top_5_journals
```




## Analytical Query II: Number of articles per journal per year broken down by quarter. Again, think of a good way to show this. Is a table better than a visualization or is it better to use a visualization.
```{r, journal_per_year, eval = T, warning = F}
# Store the result to a dataframe
journal_per_year <- dbGetQuery(mysqlCon, 
   "SELECT jd.title AS journal_title, jf.year, jf.quarter, SUM(jf.articles_count) AS sumPerQuarter
    FROM Journal_Facts jf JOIN Journal_Dim jd ON jf.journal_id = jd.journal_id
    GROUP BY jd.title, jf.year, jf.quarter")
journal_per_year
```
```{r}
ggplot(data = journal_per_year, aes(x = year, y = sumPerQuarter, fill = factor(quarter))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ journal_title) +
  labs(title = "Number of Articles per Journal per Year by Quarter",
       x = "Year",
       y = "Number of Articles",
       fill = "Quarter") +
  theme_minimal()

```


# Conclusion

In this report, we presented the results of two analytical queries on our MySQL data warehouse:

1. We identified the top five journals with the most articles published in them for the time period. This information can help analysts and managers understand which journals are the most popular and influential in their field.
2. We analyzed the number of articles per journal per year, broken down by quarter. This data provides insight into the publishing patterns of different journals, allowing analysts and managers to track trends and plan for future publication needs.

By presenting the data in both tabular and visual formats, we have made it easy for users to quickly understand and interpret the results of our analysis.

